---
title: "Mahalanobis model: occurrence data only"
author: "Laura Jiménez and Carola Franzen"
date: "July 2021"
output: html_document
fontsize: 12pt
linkcolor: dodgerblue4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning=FALSE,
  echo = TRUE)

```
\

# Mahalanobis distance

<!-- description -->
\


### Input and Output

In order to calculate the Mahalanobis distance and Mahalanobis model, mu and Sigma (covariance) of variables are necessary. This could, for example be the environmental data that has been correlated to the geographic occurrence data of a species.

The output can be further processed into an ellipse or a suitability map.
\

### Dependencies
**Functions:** `dismo`, `raster`

**Examples:** `rgdal`, `tools`, `sp`
\
\

## Worked Examples

### Read Source Code and Libraries 

```{r}
source(".\\Functions\\TgetEcoord.R")
source(".\\Functions\\TnicheG.R")
```

```{r}
library(dismo)
library(raster)
library(rgdal)
library(tools)
library(sp)
```
\


## *Threnetes ruckeri* - example with two environmental layers
\

**Input Data**

- Threnetes_ruckeri_occ_GE.csv
\
\

Read table with environmental data that are the basis for the calculation of the Mahalanobis distance.
```{r}
thr.occ_E <- read.csv("./Generated_Data/Threnetes_ruckeri_occ_GE.csv",header=T)[,-(1:2)]
```
\

The centroid is calculated by taking the means of the two columns (`mu`) in the matrix with two environmental layers. This parameter will set the values of the maximum likelihood estimate.
```{r}
(center <- colMeans(thr.occ_E))
```
\

Calculate the covariance of the environmental data (`Sigma`).
```{r}
(boundary <- cov(thr.occ_E))
```
\

Here, add example on how to plot the principal axis of the estimated ellipse (use this example: \url{https://multivariate-ecology.netlify.app/lab-5-transformations-eigenvectors-and-eigenvalues.html})

\

\

## *Catasticta nimbice*

### 1) Get environmental information from three raster layers
\

**Input Data**

- Catasticta_nimbice_occ_G.csv
- cn_bio1_cropped.tif
- cn_bio6_cropped.tif
- cn_bio12_cropped.tif
\
\

Read necessary files from the Generated_Data folder for *Catasticta nimbice*: a matrix with coordinates of species observations (latitude and longitude) and the three (cropped) raster layers for bio1, bio6, and bio12.  
```{r, results='hide'}
# read occurrence data
occ_G <- read.csv("./Initial_Data/Catasticta_nimbice_occ_G.csv",header=T)[,-1]

# read raster files
bio1cn <- raster("./Generated_Data/cn_bio1_cropped.tif")
bio6cn <- raster("./Generated_Data/cn_bio6_cropped.tif")
bio12cn <- raster("./Generated_Data/cn_bio12_cropped.tif")


```
\

Prepare the parameters: create a rasterstack of the cropped environmental layers and create a vector with the environmental layer names.
```{r}
# stack the layers
bios <- stack(bio1cn, bio6cn, bio12cn)
# create vector with names in same order as stack
E.names <- c("bio1", "bio6", "bio12")

```
\

Apply function
```{r}
cn.GE3 <- get.Ecoord(Estck = bios, Gcoord = occ_G, Enames = E.names)
head(cn.GE3, n=4)
```
\

Write as csv
```{r}
write.csv(cn.GE3,file=paste0("./Generated_Data/Catasticta_nimbice_occ_GE3.csv"),
          row.names = F)
```
\

**Output Data**

- Catasticta_nimbice_occ_GE3.csv
\
\


### 2) Calculate suitability index
\

**Input Data**

- Catasticta_nimbice_occ_GE3.csv
\

Prepare parameters for the function `niche.G`. For this, a table with only the variables is necessary.
```{r}
occ_E <- cn.GE3[,-(1:2)]
```
\

The centroid is calculated by taking the means of the two columns in the matrix (`mu`). This parameter will set the values of the maximum likelhood estimate.
```{r}
center <- colMeans(occ_E)
```
\

Calculate the covariance of the environmental data (`Sigma`).
```{r}
boundary <- cov(occ_E)
```
\

Apply function
```{r}
cn.maha <- niche.G(Estck = bios, mu = center, Sigma = boundary)
```
\

Save resulting raster. This file will be necessary for further calculations.
```{r}
writeRaster(cn.maha, "./Generated_Data/cn_maha3_cropped.tif", overwrite= T) 
```
\

Plot.
```{r}
plot(cn.maha)
```
\

**Output Data**

- Catasticta_nimbice_maha3_cropped.tif
\
\

### 3) Comparing suitability index against Mahalanobis distance
\

The Mahalanobis distance and the suitability index calculated with the Mahalanobis model are slightly different. If applied to the raster cells, the patterns might look the same but the scale of the values will look different. 

\

**Input Data**

- Catasticta_nimbice_maha3_cropped.tif
\
\

Calculate the Mahalanobis distance and apply it to the cells.
```{r}
maha.fun <- function(cell){
    X <- as.vector(cell, mode = "numeric")
    sui.ind <- mahalanobis(x= X, center= center, cov= boundary)
    return(sui.ind)
}

maha.map <- calc(bios,fun=maha.fun)
```
\

If applied to the raster cells, the patterns will look the same but the scale of the values will look different:
```{r}
mahacompare <- stack(maha.map, cn.maha)

plot(mahacompare)
```


