---
title: "Mahalanobis distance"
author: "Laura Jimenéz and Carola Franzen"
date: "July 2021"
output: html_document
fontsize: 12pt
linkcolor: dodgerblue4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dismo)
library(raster)
library(rgdal)
library(rgeos)
library(tools)
library(sp)
library(ggpubr)

```

# Mahalanobis distance

## *Catasticta nimbice*

1: get.Ecoord -> occ_GE3.csv
2: rs.inE -> get random sample points
3: fit.Niche -> get mahalanobis calculation
4: nicheG -> get raster with model
5: apply accu curve?

### Get environmental information from three raster layers

Read necessary files from the Initial_Data folder for *Catasticta nimbice*: a matrix with coordinates of species observations (latitude and longitude) and the three raster layers for bio1, bio6, and bio12.  
```{r, results='hide'}
# read occurrence data
occ_G <- read.csv("./Initial_Data/Catasticta_nimbice_occ_G.csv",header=T)[,-1]

# read raster files
bio1 <- raster("./Initial_Data/bio1WH.asc")
bio6 <- raster("./Initial_Data/Bio6_america.tif")
bio12 <- raster("./Initial_Data/bio12WH.asc")

# read shapefile of study area
cn.shp <- readOGR("./Initial_Data/shapefiles","nimbice3")
```
\

Prepare the parameters: crop all rasters by the study region (shapefile) and then create a rasterstack of it, and create a vector with the environmental layer names.
```{r}
bio1c <- mask(crop(bio1, cn.shp), cn.shp)
bio12c <- mask(crop(bio12, cn.shp), cn.shp)
bio6c <- mask(crop(bio6, cn.shp), cn.shp)
# the extent of bio6 does not match the other layers -> values need to be adjusted
bio6c.ext <- raster(vals=values(bio6c),ext=extent(bio1c),crs=crs(bio1c),
              nrows=dim(bio1c)[1],ncols=dim(bio1c)[2])
# stack the layers
bios <- stack(bio1c, bio6c.ext, bio12c)
# create vector with names in same order as stack
E.names <- c("bio1", "bio6", "bio12")

```
\

Apply source code
```{r}
source(".\\Functions\\TgetEcoord.R")
```
\

Apply function
```{r}
cn.GE3 <- get.Ecoord(Estck = bios, Gcoord = occ_G, Enames = E.names)
head(cn.GE3, n=4)
```
\

Write as csv
```{r}
write.csv(cn.GE3,file=paste0("./Generated_Data/Catasticta_nimbice_occ_GE3.csv"),
          row.names = F)
```
\

\
<!---- put mean and covariance _OR_ do the whole fitniche calculation ---->

### Calculate Mahalanobis distance

Prepare parameters for the function `niche.G`

```{r}
occ_E <- cn.GE3[,-(1:2)]

center <- colMeans(occ_E)

boundary <- cov(occ_E)
```
\

Apply source code for `niche.G`
```{r}
source(".\\Functions\\TnicheG.R")
```

Apply function
```{r}
#cn.maha <- niche.G(Estck = bios, mu = center, Sigma = boundary)
```
\

<!---- _OR_ do the whole fitniche calculation ---->


### Produce random samples of the study area for *Catasticta nimbice*

Apply source code
```{r}
source(".\\Functions\\Trandom_sampling_inE.R")
```
\

Apply the function `rs.inE` with the previously loaded and prepared parameters and decide on a sampling size.
```{r}
cn.sample <- rs.inE(region = cn.shp, N = 8000, Estck = bios)
head(cn.sample, n=3)
tail(cn.sample, n=3)
```
\

### Create values for the Mahalanobis distance with fit.niche

Apply source code
```{r}
source(".\\Functions\\Tfit_wn_maha_model.R")
```
\

Apply function
```{r}
#cn.maha2 <- fitNiche(E.occ = occ_E, E.samM = cn.sample)
```

