---
title: "Mahalanobis distance"
author: "Laura Jimenéz and Carola Franzen"
date: "July 2021"
output: html_document
fontsize: 12pt
linkcolor: dodgerblue4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dismo)
library(raster)
library(rgdal)
library(rgeos)
library(tools)
library(sp)
library(ggpubr)

```

# Mahalanobis distance


## *Threnetes ruckeri* - example with two environmental layers

Read table with environmental data that are the basis for the calculation of the Mahalanobis distance.
```{r}
thr.occ_E <- read.csv("./Generated_Data/Threnetes_ruckeri_occ_GE.csv",header=T)[,-(1:2)]
```
\

The centroid is calculated by taking the means of the two columns (`mu`) in the matrix with two environmental layers. This parameter will set the values of the maximum likelihood estimate.
```{r}
(center <- colMeans(thr.occ_E))
```
\

Calculate the covariance of the environmental data (`Sigma`).
```{r}
(boundary <- cov(thr.occ_E))
```
\

Here, add example on how to plot the principal axis of the estimated ellipse (use this example: \url{https://multivariate-ecology.netlify.app/lab-5-transformations-eigenvectors-and-eigenvalues.html})

\

\

## *Catasticta nimbice*

1: get.Ecoord -> occ_GE3.csv
2: rs.inE -> get random sample points
3: fit.Niche -> get mahalanobis calculation
4: nicheG -> get raster with model
5: apply accu curve?

<!--- from here it is how to get occ_table with 3 rasters: GE3 --->

### Get environmental information from three raster layers

Read necessary files from the Generated_Data folder for *Catasticta nimbice*: a matrix with coordinates of species observations (latitude and longitude) and the three (cropped) raster layers for bio1, bio6, and bio12.  
```{r, results='hide'}
# read occurrence data
occ_G <- read.csv("./Initial_Data/Catasticta_nimbice_occ_G.csv",header=T)[,-1]

# read raster files
bio1cn <- raster("./Generated_Data/Catasticta_nimbice_bio1_cropped.tif")
bio6cn <- raster("./Generated_Data/Catasticta_nimbice_bio6_cropped.tif")
bio12cn <- raster("./Generated_Data/Catasticta_nimbice_bio12_cropped.tif")


```
\

Prepare the parameters: create a rasterstack of the cropped environmental layers and create a vector with the environmental layer names.
```{r}
# stack the layers
bios <- stack(bio1cn, bio6cn, bio12cn)
# create vector with names in same order as stack
E.names <- c("bio1", "bio6", "bio12")

```
\

Apply source code
```{r}
source(".\\Functions\\TgetEcoord.R")
```
\

Apply function
```{r}
cn.GE3 <- get.Ecoord(Estck = bios, Gcoord = occ_G, Enames = E.names)
head(cn.GE3, n=4)
```
\

Write as csv
```{r}
write.csv(cn.GE3,file=paste0("./Generated_Data/Catasticta_nimbice_occ_GE3.csv"),
          row.names = F)
```
\

\
<!---- from here it is niche.G ---->

### Calculate Mahalanobis distance

Prepare parameters for the function `niche.G`. For this, a table with only the variables is necessary.

```{r}
occ_E <- cn.GE3[,-(1:2)]
```
\

The centroid is calculated by taking the means of the two columns in the matrix (`mu`). This parameter will set the values of the maximum likelhood estimate.
```{r}
center <- colMeans(occ_E)
```
\

Calculate the covariance of the environmental data (`Sigma`).
```{r}
boundary <- cov(occ_E)
```
\


Apply source code for `niche.G`
```{r}
# source(".\\Functions\\TnicheG.R")
```
\

Apply function
```{r}
cn.maha <- niche.G(Estck = bios, mu = center, Sigma = boundary)
```
\

```{r}
writeRaster(cn.maha, "./Generated_Data/Catasticta_nimbice_maha3_cropped.tif", overwrite= T) 
```
\

Plot
```{r}
plot(cn.maha)
```
\

### Create a table with a suitability index

