---
title: "Mahalanobis distance"
author: "Laura Jimenéz and Carola Franzen"
date: "July 2021"
output: html_document
fontsize: 12pt
linkcolor: dodgerblue4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dismo)
library(raster)
library(rgdal)
library(rgeos)
library(tools)
library(sp)
library(ggpubr)

```

# Mahalanobis distance

## *Catasticta nimbice*

1: get.Ecoord -> occ_GE3.csv
2: rs.inE -> get random sample points
3: fit.Niche -> get mahalanobis calculation
4: nicheG -> get raster with model
5: apply accu curve?

### Get environmental information from three raster layers

Read necessary files from the Generated_Data folder for *Catasticta nimbice*: a matrix with coordinates of species observations (latitude and longitude) and the three (cropped) raster layers for bio1, bio6, and bio12.  
```{r, results='hide'}
# read occurrence data
occ_G <- read.csv("./Initial_Data/Catasticta_nimbice_occ_G.csv",header=T)[,-1]

# read raster files
bio1cn <- raster("./Generated_Data/Catasticta_nimbice_bio1_cropped.tif")
bio6cn <- raster("./Generated_Data/Catasticta_nimbice_bio6_cropped.tif")
bio12cn <- raster("./Generated_Data/Catasticta_nimbice_bio12_cropped.tif")


```
\

Prepare the parameters: create a rasterstack of the cropped environmental layers and create a vector with the environmental layer names.
```{r}
# stack the layers
bios <- stack(bio1cn, bio6cn, bio12cn)
# create vector with names in same order as stack
E.names <- c("bio1", "bio6", "bio12")

```
\

Apply source code
```{r}
source(".\\Functions\\TgetEcoord.R")
```
\

Apply function
```{r}
cn.GE3 <- get.Ecoord(Estck = bios, Gcoord = occ_G, Enames = E.names)
head(cn.GE3, n=4)
```
\

Write as csv
```{r}
write.csv(cn.GE3,file=paste0("./Generated_Data/Catasticta_nimbice_occ_GE3.csv"),
          row.names = F)
```
\

\
<!---- put mean and covariance _OR_ do the whole fitniche calculation ---->

### Calculate Mahalanobis distance

Prepare parameters for the function `niche.G`

```{r}
occ_E <- cn.GE3[,-(1:2)]

center <- colMeans(occ_E)

boundary <- cov(occ_E)
```
\

Apply source code for `niche.G`
```{r}
source(".\\Functions\\TnicheG.R")
```
\

Apply function
```{r}
cn.maha <- niche.G(Estck = bios, mu = center, Sigma = boundary)
```
\

```{r}
writeRaster(cn.maha, "./Generated_Data/Catasticta_nimbice_maha3_cropped.tif", overwrite= T) 
```
\

Plot
```{r}
plot(cn.maha)
```
\

### Create a table with a suitability index

Prepare parameters: the file for `G.occ` has already been loaded, but for the parameter `suit.Estck` the rasters of the model and environmental layers need to be stacked first. 
```{r}
cnmaha3 <- raster("./Generated_Data/Catasticta_nimbice_maha3_cropped.tif")

maha.stack <- stack(cnmaha3, bio1cn, bio6cn, bio12cn)
```
\

Load the source:
```{r}
source(".\\Functions\\Tgettable.R")
```
\

Apply the function:
```{r}
table.cnmaha <- get.table(G.occ = occ_G, suit.Estck = maha.stack)
```

