---
title: "Function *sam.polyM*"
author: "Laura Jimenéz and Carola Franzen"
date: "June 2021"
output: html_document
fontsize: 12pt
linkcolor: dodgerblue4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(rgdal)
library(rgeos)
library(tools)
```
\

## Sampling random points from an area of interest

The function `sam.polyM` takes a random sample of environmental combinations inside the study area. For this, the environmental data from a rasterstack is clipped by a shapefile that delimits the area before the sample is taken.
\

\

### Parameters

* `M.shp` = a shapefile of the study area (polygon)
\

* `N` = the sample size
\

* `bios` = a rasterstack that contains at least two layers with environmental data
\

\

### Dependencies

* `raster`
\

* `rgdal`
\

* `rgeos`
\

* `tools`
\

\

## Function's code


### How *sam.polyM* works

The function `sam.polyM`uses a shapefile of the area of interest as a mask to clip the rasterstack in order to reduce the information of the rasterstack to the same geographical area. As a next step, cells with an NA value will be deleted from the clipped rasterstack. As a next step, a random sample of indices is generated. With these indices a sample of the clipped rasterstack is taken. The sample size is dictated by N.

```{r}
# Get a random sample of points inside the polygon that delimits M (= )
# and extract their environmental values
sam.polyM <- function(M.shp,N,bios){
  # crop and mask the environmental layers with the M polygon
  clip.M <- mask(crop(bios,M.shp),M.shp)
  # get rid of cells with NA values = indices
  ind <- which(!is.na(clip.M[[1]][]))
  # get a random sample of indices
  sam <- sample(ind,N,replace = T)
  # choose the points corresponding to the selected indices
  Mpnts <- clip.M[][sam,]
  return(Mpnts)
}
```
\

### Output

The output of the function is a matrix of random points with as many columns as layers the rasterstack has.
\

\

## Worked Examples

### Input Data

The raster files need to contain information on climatic values that are of interest for the user. The rasters used in the examples below are from [WorldClim](https://worldclim.org/data/index.html) and have a resolution of "10 minutes". The files usually do not need further editing.

A polygon that marks the borders of a species' occurrence or the study area is necessary. This polygon should be a shapefile.
\

### *Catasticta nimbice*

First, the rasterstack for the parameter `bios` is prepared. For this example two rasters with the environmental data "annual mean temperature" and "total annual precipitation" are used. The rasterstack will therefore have two layers.

```{r}
# Read environmental layers
bio1 <- raster("./Input_Data/bio1WH.asc")
bio12 <- raster("./Input_Data/bio12WH.asc")

# Create a single raster with as many layers as environmental variables
stck_bios <- stack(bio1, bio12)
```
\

Read the parameter for `M.shp`.

```{r, results='hide', message=FALSE}
M.shp <- readOGR("./Input_Data/shapefiles","nimbice3")
```
\

Apply the function with `bios`, `M.shp`, and `N`. Here, `N` will be 10,000. The generated matrix will therefore have two columns with environmental data from the study area and 10,000 rows.

```{r}
sam.Mpnts <- sam.polyM(M.shp = M.shp,N = 10000,bios = stck_bios)
```
\

Save the output as a csv-file (do not run everytime)

```{r, eval=FALSE}
write.csv(sam.Mpnts,file=paste0("./Output_Data/Catasticta_nimbice","_M_bios.csv"),
          row.names = F)
```
\

Create a simple plot of the environmental space (E-Space) with a species' occurrence.
```{r, fig.show='hide'}
# read occurrence data
occ <- read.csv("./Input_Data/Catasticta_nimbice_bios.csv",header=T) 
# set colors
pal <- c("grey50", "turquoise")

# plot
x11()
#   use random points with environmental data for the plot
plot(sam.Mpnts[,1], sam.Mpnts[,2], pch=".", col=pal[1], 
     xlab="Mean Annual Temperature", ylab="Accumulated Precipitation", 
     main="Environmental Space") 
# add points of species environmental data
points(occ$bio1, occ$bio12, pch=19, col=pal[2])
# add legend
legend(x= "topleft",
       legend = c("Study Area", "Occurences"),
       pch = c(20, 19),
       col = c(pal[1], pal[2]),
       bty = "n")
```
\



### *Threnetes ruckeri*

The rasterstack is the same as the previous example. In this example, a different graph will be plotted using the sampled points.

```{r}
# Read M polygon
M.shp <- readOGR("./threnetes_shp","Threnetes_ruckeri") ###

## Get a random sample of points in M and extract its corresponding environmental values
sam.Mpnts <- sam.polyM(M.shp = M.shp,N = 10000,bios = stck_bios)


```


